---
title: "HW4"
format: 
 html:
   embed-resources: true
---

## Import the library

```{r}
library(nycflights13)
library(tidyverse)
library(lubridate)
library(ggplot2)
```

## Problem 1 - Tidyverse

**a**

```{r}
#check the two dataframes
flights_df <- as.data.frame(flights)
airpotts <- as.data.frame(airports)
```

Firstly, we get the table for departure_delay.

```{r}
#Departure-delay Table
dep_delay <- flights %>%
  #get the airport names
  inner_join(airports, by = c("dest" = "faa")) %>%
  group_by(name) %>%
  summarise(
    mean_dep = mean(dep_delay, na.rm = TRUE),
    median_dep = median(dep_delay, na.rm = TRUE),
    #count for flights
    flights = n()
  ) %>%
  filter(flights >= 10) %>%
  arrange(desc(mean_dep)) %>%
  # Exclude counts for flights
  select(-flights)

dep_delay
```

Then, we get the table for arrival delay.

```{r}
#Arrival-delay Table
Arr_delay <- flights %>%
  #inner join two tables
  #get the airport names
  inner_join(airports, by = c("dest" = "faa")) %>%
  group_by(name) %>%
  summarise(
    mean_arr = mean(arr_delay, na.rm = TRUE),
    median_arr = median(arr_delay, na.rm = TRUE),
    #count for flights
    flights = n()
  ) %>%
  filter(flights >= 10) %>%
  arrange(desc(mean_arr)) %>%
  # Exclude counts for flights
  select(-flights)

Arr_delay
```

**b**

```{r}
#check the dataframe
planes <- as.data.frame(planes)

fast_aircraft_model <- flights %>%
  #calculate MPH_speed
  mutate(MPH_speed = distance / (air_time / 60)) %>%
  group_by(tailnum) %>%
  summarise(
    avg_speed = mean(MPH_speed, na.rm = TRUE),
    flights = n()) %>%
  #inner join two tables
  inner_join(planes, by = c("tailnum")) %>%
  arrange(desc(avg_speed)) %>%
  slice(1) %>%
  select(model, avg_speed, flights)
  
fast_aircraft_model
```

## Problem 2 - get_temp()

```{r}
#import the csv
nnmaps <- read.csv("chicago-nmmaps.csv")

get_temp <- function(month, year, data, celcius = FALSE, average_fn = mean){
  #check for month-parameter
  if(is.character(month)){
    month_1 <- month
    month <- match(tolower(month), tolower(month.name))
    if (is.na(month)){
      month <- match(tolower(month_1), tolower(month.abb))
    }}
  if(is.na(month) | month < 1 | month > 12){
    stop("Please provide a valid month (1-12 or a character name).")
  }

  #check for year-parameter
  if(!is.numeric(year) | year <= 0){
    stop("Please provide a valid year.")
  }
  
  find_data <- data %>%
    mutate(date = ymd(date)) %>%
    filter(month(date) == as.integer(month), year(date) == year)
  
  #check if the data is empty
  if (nrow(find_data) == 0) {
    stop("No data available for the specified month and year.")
  }
  
  #calculate the average temp
  avg_temp <- find_data %>%
    summarise(avg_temp = average_fn(temp, na.rm = TRUE)) %>%
    select(avg_temp)
  
  #check if we need to change celsius
  if (celsius) {
    avg_temp <- (avg_temp - 32) * (5 / 9)
  }
  
  return(avg_temp)
}

get_temp("jan", 1997, data = nnmaps)
```

## Problem 3 - Visualization

```{r}
#import the csv
market <- read.csv("df_for_ml_improved_new_market.csv")
```

**a** Yes, there is. Based on the plot, we observe that **the sales price in USD** shows an overall upward trend over time, but there are significant fluctuations after 2007, with a peak in 2008.

```{r}
ggplot(market, aes(x = year, y = price_usd)) +
  stat_summary(fun = mean, geom = "line", color = "purple", alpha = 0.7, size = 1) +
  stat_summary(fun = mean, geom = "point", color = "blue", size = 2) +
  labs(
    title = "Average Sales Price Over Time",
    x = "Year",
    y = "Average USD Sales Price"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 10)
  )
```

**b** Yes. Based on the plot, we can observe that the sales proportion of the Painting category is declining. The Photograph category saw a significant increase in 1999, followed by a stable trend. The Print category started appearing in 2000, with a relatively small but steady share. The Sculpture category reached its lowest proportion in 1999, then increased and stabilized afterward.

```{r}
#get a new dataframe
#including years and the counts of items of different genres
genre_columns <- grep("Genre___", names(market), value = TRUE)
genre_data <- market %>%
  select(year, all_of(genre_columns)) %>%
  group_by(year) %>%
  summarise(across(everything(), sum))

#transpose the dataframe
#get the genre-percentage
genre_1 <- genre_data %>%
  pivot_longer(cols = -year, names_to = "Genre", values_to = "Count") %>%
  group_by(year) %>%
  mutate(Percentage = Count / sum(Count) * 100) %>%
  mutate(Genre = gsub("Genre___", "", Genre))

ggplot(genre_1, aes(x = factor(year), y = Percentage, fill = Genre)) +
  geom_bar(stat = "identity") +
  labs(title = "Distribution of Genre of Art Sales Over Years", x = "Year", y = "Percentage") +
  theme_minimal() +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  scale_fill_brewer(palette = "Pastel1") +
  theme(legend.position = "top", 
        legend.title = element_text(size = 10),
        plot.title = element_text(hjust = 0.5))
```

**c**Observing the plot, we can see that the overall trend in art sales prices aligns closely with the trends in sales prices across different categories. The Photograph category is the primary influencing factor on the overall trend, while the price trend of Print artworks often contrasts with the overall price trend.

```{r}
#get a new dataframe
#including years and the counts of items of different genres
genre_2 <- market %>%
  pivot_longer(cols = starts_with("Genre___"), names_to = "Genre", values_to = "Is_Genre") %>%
  filter(Is_Genre == 1) %>%
  mutate(Genre = gsub("Genre___", "", Genre))

#calculate average USD price based on genre
genre_price <- genre_2 %>%
  group_by(year, Genre) %>%
  summarise(avg_price = mean(price_usd, na.rm = TRUE))

ggplot(genre_price, aes(x = year, y = avg_price, color = Genre)) +
  geom_line(size = 1) +
  geom_point() +
  labs(title = "Sales Price Over Time by Genre", x = "Year", y = "Average Sales Price") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "top", 
        legend.title = element_text(size = 10)) +
  scale_color_brewer(palette = "Pastel2")
```
