---
title: "HW6"
format: 
 html:
  embed-resources: true
---

## Github

https://github.com/yimingliu0117/STATS506

## Import the library

```{r}
library(DBI)
library(RSQLite)
library(dplyr)
library(parallel)
library(future)
library(furrr)
```

## Stratified Bootstrapping

**a**

```{r}
#import SQL data
lahman <- dbConnect(RSQLite::SQLite(), "lahman_1871-2022.sqlite")
fielding_data <- dbReadTable(lahman, "Fielding")

#calculate the RF value
fielding_data <- fielding_data %>%
  filter(InnOuts > 0, !is.na(InnOuts)) %>%
  mutate(RF = 3 * (PO + A) / InnOuts) %>%
  filter(!is.na(RF))  # Remove NA data

#calculate the average RF for each team
team_rf <- fielding_data %>%
  group_by(teamID) %>%
  summarise(avg_RF = mean(RF, na.rm = TRUE)) %>%
  arrange(desc(avg_RF))

#get Top10 team
top_teams <- team_rf %>%
  slice_max(avg_RF, n = 10) %>%
  pull(teamID)

fielding_top <- fielding_data %>%
  filter(teamID %in% top_teams)

```

Let's check about TOP10 teams.
```{r}
unique(fielding_top$teamID)
```

```{r}
#' Stratified_bootstrap
#'
#' @param data Input parameter
#' @param n_boot Default argument specifying the number of bootstrap samples to generate 
#'
#' @return The function returns a data frame containing the computed statistics for all bootstrap iterations.
#'
stratified_bootstrap <- function(data, n_boot = 1000) {
  team_ids <- unique(data$teamID)
  boot_results <- replicate(n_boot, {
    sampled_data <- lapply(team_ids, function(team) {
      team_data <- data %>% filter(teamID == team)
      team_data[sample(nrow(team_data), replace = TRUE), ]
    })
    boot_sample <- do.call(rbind, sampled_data)
    boot_sample %>%
      group_by(teamID) %>%
      summarise(avg_RF = mean(RF, na.rm = TRUE)) %>%
      arrange(desc(avg_RF))
  }, simplify = FALSE)
    boot_results_df <- bind_rows(boot_results, .id = "bootstrap")
  return(boot_results_df)
}
```

1.Without any parallel processing

```{r}
set.seed(42)
boot_1 <- stratified_bootstrap(fielding_top)
print(boot_1, n = Inf)
```

2.Using parallel processing with the parallel package

```{r}
set.seed(42)

#export necessary objects and load required packages in workers
#load dplyr in each cluster
cl <- makeCluster(detectCores() - 1)
clusterExport(cl, c("fielding_top", "stratified_bootstrap"))
clusterEvalQ(cl, library(dplyr))

#perform parallel bootstrap
boot_2 <- parLapply(cl, 1:1000, function(i) {
  stratified_bootstrap(fielding_top, n_boot = 1)
})

stopCluster(cl)

#combine results
boot_2 <- bind_rows(boot_2, .id = "bootstrap")
print(boot_2, n = Inf)
```

3.Using futures with the future package

```{r}
set.seed(42)

#use multisession plan with parallel workers
plan(multisession, workers = detectCores() - 1)

# **future_map_dfr** applies a function to each element of a vector in parallel and combines the results into a single data frame by row
boot_3 <- future_map_dfr(
  1:1000,
  ~ {
    stratified_bootstrap(fielding_top, n_boot = 1)
  },
  .options = furrr_options(seed = TRUE)  #specify seed = TRUE for parallel-safe RNG
)

#add some functions to avoid warnings
#resets the future plan to avoid lingering parallel workers.
plan(sequential)
#clears unused resources.
gc()
print(boot_3, n = Inf)
```

**b**

```{r}
#' define the standard errors
#'
#' @param boot_results 
#'
#' @return descending average RF for Top10 teams
#'
#' @examples
#' results_1 <- calculate_se(boot_1)
#' results_2 <- calculate_se(boot_2)
#' results_3 <- calculate_se(boot_3)
#' 
calculate_se <- function(boot_results) {
  boot_results %>%
    group_by(teamID) %>%
    summarise(
      aveg_RF = mean(avg_RF),
      se_RF = sd(avg_RF)
    ) %>%
    filter(teamID %in% top_teams) %>%
    arrange(desc(aveg_RF))
}

results_1 <- calculate_se(boot_1)
results_2 <- calculate_se(boot_2)
results_3 <- calculate_se(boot_3)

#combine the results
final_results <- bind_rows(
  results_1 %>% mutate(method = "Serial"),
  results_2 %>% mutate(method = "Parallel"),
  results_3 %>% mutate(method = "Future")
)

print(final_results, n = Inf)
```

**c**

```{r}
#calculate 
time_1 <- system.time({
  boot_1 <- stratified_bootstrap(fielding_top, n_boot = 1000)
})

# 并行计算（parallel 包）的时间
time_2 <- system.time({
  cl <- makeCluster(detectCores() - 1) 
  clusterExport(cl, c("fielding_top", "stratified_bootstrap")) 
  clusterEvalQ(cl, library(dplyr)) 
  
  boot_2 <- parLapply(cl, 1:1000, function(i) {
    stratified_bootstrap(fielding_top, n_boot = 1)
  })
  
  stopCluster(cl)
  boot_2 <- bind_rows(boot_2, .id = "bootstrap") # 合并结果
})

time_3 <- system.time({
  plan(multisession, workers = detectCores() - 1) # 设置多进程
  boot_3 <- future_map_dfr(
    1:1000,
    ~ {
      stratified_bootstrap(fielding_top, n_boot = 1)
    },
    .options = furrr_options(seed = TRUE)
  )
  plan(sequential)
  gc()
})


time_results <- data.frame(
  Method = c("Serial", "Parallel", "Future"),
  User_Time = c(time_1["user.self"], time_2["user.self"], time_3["user.self"]),
  System_Time = c(time_1["sys.self"], time_2["sys.self"], time_3["sys.self"]),
  Elapsed_Time = c(time_1["elapsed"], time_2["elapsed"], time_3["elapsed"])
)

print(time_results)
```

**Serial execution takes the longest time**: For small-scale tasks, the serial method is efficient enough, but as the task size increases, the time cost grows significantly.

**Parallel and Future methods are faster**: As the number of cores increases, parallel efficiency improves, but the overhead from inter-process communication may also grow.

**Parallel is slightly better than Future**: The future method incurs slight additional overhead from managing task scheduling.
