---
title: "HW1"
format: html
---

## Problem 1: Wine Data

**(a)** Firstly,read data file and make it to be a dataframe. Then use names() function to give the column names.

```{r}
#getwd() --check about the working destination
winedata <- read.table("wine.data", sep = ",")
df1 <- data.frame(winedata)

names(df1) <- c("Class", "Alcohol", "Malic_acid", "Ash", "Alcalinity_of_ash", "Magnesium", "Total_phenols", "Flavanoids", "Nonflavanoid_phenols", "Proanthocyanins", "Color_intensity", "Hue", "Diluted_wines", "Proline")
df1
```

**(b)** Use table() function to calculate class numbers and do comparison.

```{r}
#count the number according to classes
count_class_df1 <- table(df1$Class)

#check the data and mode
count_class_df1
mode(count_class_df1)

#input the number from wine.names file
count_class_winemames <- c("Class 1" = 59, "Class2" = 71, "class3" = 48)

#check the data and mode
count_class_winemames
mode(count_class_winemames)

#check if they are the same
all(count_class_df1 == count_class_winemames)
                                    
```

**(c1)** The correlation between alcohol content and color intensity is 0.5463642.

```{r}
#use cor() function to calculate 
correlation_alcohol_color <- cor(df1$Alcohol,df1$'Color_intensity') 
correlation_alcohol_color
```

**(c2)** The highest is class1 and the lowest is class2.

```{r}
#split the data into 3 different dataframes for the next step
#learn the split() function from google
wine_class <- split(df1, df1$Class)
wine_class

#use sapply() function to calculate different correlations
correlations_class <- sapply(wine_class, function(df) {
  cor(df$Alcohol, df$Color_intensity)
})
correlations_class

#learn which.max() function from google
Highest_Cor <- names(which.max(correlations_class))
Lowest_Cor <- names(which.min(correlations_class))
Highest_Cor
Lowest_Cor
```

**(c3)** The alcohol content is 14.34.

```{r}
#find the max color_intensity row
highest_color_inten <- df1[which.max(df1$Color_intensity), ]
highest_color_inten

#get the alcohol content
highest_color_inten$Alcohol

```

**(c4)** It's 8.426966%

```{r}
#get the number of qualified wine and the number of total
num_higher_proanth <- sum(df1$Proanthocyanins > df1$Ash)
num_ttl <- dim(df1)[1]

#get the percentage
(num_higher_proanth/num_ttl)*100
```

**(d)** Get different part and combine them.

```{r}
#get overall averages and class-averages
overall_averages <- colMeans(df1[,-1])
wine_class <- split(df1, df1$Class)

class_averages <- sapply(wine_class, function(df) {
  colMeans(df[,-1])
})

#use t() function to reverse the row and column
#use rbind() function to combine dataframes
#learn those functions from google
class_averages <- t(class_averages)
averages_table <- rbind("Overall" = overall_averages, class_averages)
averages_table
```

**(e)** Class1 has significantly higher phenol levels compared to Class2 and Class3.Class2 has significantly higher phenol levels compared to Class3.

```{r}
#define the t-test function
wine_t_tests <- function(class1, class2) {
  set1 <- subset(df1, Class == class1)$Total_phenols
  set2 <- subset(df1, Class == class2)$Total_phenols
  t.test(set1, set2)
}

#the test process
list("class1_2" = wine_t_tests(1, 2),
     "Class1_3" = wine_t_tests(1, 3),
     "Class2_3" = wine_t_tests(2, 3)
)
```

## Problem 2: AskAManager

**(a)** Import the Data

```{r}
Askamanager <- read.table("AskAManager.csv", sep = ",")
df2 <- data.frame(Askamanager)
df2
```

**(b)**
```{r}
names(df2) <- c()
```





## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
