---
title: "HW1_Yiming Liu"
format: html
---

## Github link https://github.com/yimingliu0117/STATS506
## Problem 1: Wine Data

**(a)** Firstly,read data file and make it to be a dataframe. Then use names() function to give the column names.

```{r}
#getwd() --check about the working destination
winedata <- read.table("wine.data", sep = ",")
df1 <- data.frame(winedata)

names(df1) <- c("Class", "Alcohol", "Malic_acid", "Ash", "Alcalinity_of_ash", "Magnesium", "Total_phenols", "Flavanoids", "Nonflavanoid_phenols", "Proanthocyanins", "Color_intensity", "Hue", "Diluted_wines", "Proline")
```

**(b)** Use table() function to calculate class numbers and do comparison.

```{r}
#count the number according to classes
count_class_df1 <- table(df1$Class)

#check the data and mode
count_class_df1
mode(count_class_df1)

#input the number from wine.names file
count_class_winemames <- c("Class 1" = 59, "Class2" = 71, "class3" = 48)

#check the data and mode
count_class_winemames
mode(count_class_winemames)

#check if they are the same
all(count_class_df1 == count_class_winemames)
                                    
```

**(c1)** The correlation between alcohol content and color intensity is 0.5463642.

```{r}
#use cor() function to calculate 
correlation_alcohol_color <- cor(df1$Alcohol,df1$'Color_intensity') 
correlation_alcohol_color
```

**(c2)** The highest is class1 and the lowest is class2.

```{r}
#split the data into 3 different dataframes for the next step
#learn the split() function from google
wine_class <- split(df1, df1$Class)

#use sapply() function to calculate different correlations
correlations_class <- sapply(wine_class, function(df) {
  cor(df$Alcohol, df$Color_intensity)
})
correlations_class

#learn which.max() function from google
Highest_Cor <- names(which.max(correlations_class))
Lowest_Cor <- names(which.min(correlations_class))
Highest_Cor
Lowest_Cor
```

**(c3)** The alcohol content is 14.34.

```{r}
#find the max color_intensity row
highest_color_inten <- df1[which.max(df1$Color_intensity), ]

#get the alcohol content
highest_color_inten$Alcohol
```

**(c4)** It's 8.426966%

```{r}
#get the number of qualified wine and the number of total
num_higher_proanth <- sum(df1$Proanthocyanins > df1$Ash)
num_ttl <- dim(df1)[1]

#get the percentage
(num_higher_proanth/num_ttl)*100
```

**(d)** Get different part and combine them.

```{r}
#get overall averages and class-averages
overall_averages <- colMeans(df1[,-1])
wine_class <- split(df1, df1$Class)

class_averages <- sapply(wine_class, function(df) {
  colMeans(df[,-1])
})

#use t() function to reverse the row and column
#use rbind() function to combine dataframes
#learn those functions from google
class_averages <- t(class_averages)
averages_table <- rbind("Overall" = overall_averages, class_averages)
averages_table
```

**(e)** Class1 has significantly higher phenol levels compared to Class2 and Class3.Class2 has significantly higher phenol levels compared to Class3.

```{r}
#define the t-test function
wine_t_tests <- function(class1, class2) {
  set1 <- subset(df1, Class == class1)$Total_phenols
  set2 <- subset(df1, Class == class2)$Total_phenols
  t.test(set1, set2)
}

#the test process
list("class1_2" = wine_t_tests(1, 2),
     "Class1_3" = wine_t_tests(1, 3),
     "Class2_3" = wine_t_tests(2, 3)
)
```

## Problem 2: AskAManager

**(a)** Import the Data

```{r}
Askamanager <- read.table("AskAManager.csv", sep = ",")
df2 <- data.frame(Askamanager)
```

**(b)** Use NULL to clear all variable names,then put in new names.

```{r}
#use NULL to clear all variable names
names(df2) <- NULL

#give simple names
names(df2) <- c("Count", "Time", "Old", "Industry", "Job", "Add_Title", "Annual_Salary", "Add_Monetary", "Currency", "Other Currency", "Add_income_context", "Country", "State", "City", "Years_overall", "Years_field", "Education_lev", "Gender", "Race")
```

**(c)** The row number of the raw data is 28062, and that of the filtered data is 23374.

```{r}
#confirming the number of observations
number_observations <- dim(df2)[1]
number_observations

#Restrict the data to those being paid in US dollars (USD)
usd_df <- subset(df2, Currency == "USD")
number_usd <- dim(usd_df)[1]
```

**(d)** Find people whose age is umder 18 and delete them.

```{r}
#confirming the number of observations
number_observations <- dim(df2)[1]
number_observations

#delete people under 18
df_above_18 <- df2[df2$Old != "under 18",]

dim(df_above_18)[1]
```

**(e)** The new sample size is 26837, and the stats information is showed.

```{r}
# Use summary() function to get the original information
summary(df2$Annual_Salary)
dim(df2)[1]

# filter unreal data and get the new information
threhold_df2 <- subset(df2, Annual_Salary < 200000 & Annual_Salary > 1000)
summary(threhold_df2$Annual_Salary)

dim(threhold_df2)[1]
```

## Problem 3: Palindromic Numbers

**a** isPalindromic

```{r}
#' isPalindromic
#'
#' @param a A positive integer
#'
#' @return A list with two elements:
#' \item{isPalindromic}{A logical value indicating if the input is palindromic.}
#' \item{reversed}{The input number with its digits reversed, as a character string.}
#'
#' @examples 
#' isPalindromic(728827)
#' isPalindromic(39951)
isPalindromic <- function(a){
    #make sure the input must be a positive integer
    if (!is.numeric(a) | a <= 0 | a != as.integer(a)) {
    stop("Input must be a positive integer.")
    }
  
    #make a to a string and cut it into pieces
    a_str <- as.character(a)
    split_a <- strsplit(as.character(a), NULL)
    
    #reverse the pieces and combine the string
    rev_a <- rev((split_a)[[1]])
    reversed_a <- paste(rev_a, collapse = "")
  
    # Check if the original and reversed strings are the same
    is_palindromic <- (a_str == reversed_a)

    # return the list
    return(list(isPalindromic = is_palindromic,
                reversed = reversed_a))
}

isPalindromic(728827)
isPalindromic(39951)
```

**b** nextPalindrome

```{r}
#' nextPalindrome
#'
#' @param a A positive integer
#'
#' @return A vector of length 1 with the next palindromic number greater than the input
#'
#' @examples
#' nextPalindrome(7152)
#' nextPalindrome(765431537)
nextPalindrome <- function(a){
    #make sure the input must be a positive integer
    if (!is.numeric(a) | a <= 0 | a != as.integer(a)) {
    stop("Input must be a positive integer.")
    }
  
    #build a function to determine whether the number is palindromic
    is_palindrome <- function(b){
      b_str <- as.character(b)
      split_b <- strsplit(as.character(b), NULL)
      rev_b <- rev((split_b)[[1]])
      reversed_b <- paste(rev_b, collapse = "")
      return(b_str == reversed_b)
    }
    
    #check from the next number until we get the right one
    c <- a+1
    while(!is_palindrome(c)){
      c <- c+1
    }
    return(c)
}

nextPalindrome(7152)
nextPalindrome(765431537)
```

**c** Find next palindrome

```{r}
nextPalindrome(391)
nextPalindrome(9928)
nextPalindrome(19272719)
nextPalindrome(109)
nextPalindrome(2)
```
