---
title: "HW5"
format: 
  html:
    embed-resources: true
---

## Github link

https://github.com/yimingliu0117/STATS506

## Import The Library

```{r}
library(Rcpp)
library(dplyr)
library(tidyr)
library(plotly)
library(nycflights13)
library(data.table)
sourceCpp("gcd_lcm.cpp")
```

## OOP Programming
**a**
```{r}
#define the S4 class
setClass("rational", 
         slots = c(numerator = "numeric", denominator = "numeric"))

#set the rational function
rational <- function(numerator, denominator) {
  if (!is.numeric(numerator) || !is.numeric(denominator)) {
    stop("Both numerator and denominator must be numeric.")
  }
  if (denominator == 0) stop("Denominator cannot be zero.")
  
  new("rational", numerator = as.integer(numerator), denominator = as.integer(denominator))
}

```


```{r}
# Define a validity checking function
validity <- function(object) {
  # Check for missing (NA) values in numerator or denominator
  if (any(is.na(object@numerator), is.na(object@denominator))) {
    return("Numerator and Denominator must not be empty")
  }
  
  # Ensure numerator and denominator are integers
  if (!is.integer(object@numerator) || !is.integer(object@denominator)) {
    return("Both Numerator and Denominator must be integers")
  }
  
  # Ensure denominator is not zero
  if (object@denominator == 0) {
    return("Denominator cannot be zero")
  }
  
  # Return TRUE if all checks pass
  return(TRUE)
}

# Assign the validity function to the class "rational"
setValidity("rational", validity)
```

```{r}
#set the show method
setMethod(
  "show",
  "rational",
  function(object) {
    cat(paste(object@numerator, "/", object@denominator, "\n")) # a string
  }
)
```


```{r}
#set the simplify method
#set a new S4 generic
setGeneric("simplify", function(object){
  standardGeneric("simplify")
})

setMethod("simplify", "rational", function(object) {
  common <- gcd(abs(object@numerator), abs(object@denominator))
  object@numerator <- object@numerator / common
  object@denominator <- object@denominator / common
  # Keep denominator positive
  if (object@denominator < 0L) { 
    object@numerator <- -object@numerator
    object@denominator <- -object@denominator
  }
  return(object)
})
```

```{r}
#set the quotient method
setGeneric("quotient", function(object, digits = 7){
  return(standardGeneric("quotient"))
})

setMethod("quotient", "rational", function(object, digits = 7) {
  result <- object@numerator / object@denominator
  print(round(result, digits))
  return(result)
})
```

```{r}
#Addition
setMethod(
  "+",
  signature(e1 = "rational", e2 = "rational"),
  function(e1, e2) {
    # lcm and new denominator at the same time
    lcm_denominator <- lcm(e1@denominator, e2@denominator)
    
    # new numerators
    numerator1_new <- e1@numerator * (lcm_denominator / e1@denominator)
    numerator2_new <- e2@numerator * (lcm_denominator / e2@denominator)
    # the sum of new numerators
    sum_numerator <- numerator1_new + numerator2_new
    
    # simplify the result
    simplify(rational(sum_numerator, lcm_denominator))
  }
)

#subtraction
setMethod(
  "-",
  signature(e1 = "rational", e2 = "rational"),
  function(e1, e2) {
    # lcm and new denominator at the same time
    lcm_denominator <- lcm(e1@denominator, e2@denominator)
    
    # new numerators
    numerator1_new <- e1@numerator * (lcm_denominator / e1@denominator)
    numerator2_new <- e2@numerator * (lcm_denominator / e2@denominator)
    # the difference of new numerators
    diff_numerator <- numerator1_new - numerator2_new
    
    # simplify the result
    simplify(rational(diff_numerator, lcm_denominator))
  }
)

#multiplication
setMethod(
  "*",
  signature(e1 = "rational", e2 = "rational"),
  function(e1, e2) {
    new_numerator <- e1@numerator * e2@numerator
    new_denominator <- e1@denominator * e2@denominator
    simplify(rational(new_numerator, new_denominator))
  }
)


#division
setMethod(
  "/",
  signature(e1 = "rational", e2 = "rational"),
  function(e1, e2) {
    if (e2@numerator == 0) stop("division by zero not allowed")
    new_numerator <- e1@numerator * e2@denominator
    new_denominator <- e1@denominator * e2@numerator
    simplify(rational(new_numerator, new_denominator))
  }
)
```

**b**
Fistly, we create three objects.
```{r}
r1 <- rational(24L, 6L)
r2 <- rational(7L, 230L)
r3 <- rational(0L, 4L)
```


```{r}
r1_show = tryCatch({
    show(r1)
  }, error = function(e) { paste("Error with r1:", e$message) })
```

```{r}
r3_show = tryCatch({
    show(r3)
  }, error = function(e) { paste("Error with r3:", e$message) })
```

```{r}
Add_r1_r2 = tryCatch({
    show(r1 + r2)
  }, error = function(e) { paste("Error with r1 + r2:", e$message) })
```

```{r}
Sub_r1_r2 = tryCatch({
    show(r1 - r2)
  }, error = function(e) { paste("Error with r1 - r2:", e$message) })
```

```{r}
Multi_r1_r2 = tryCatch({
    show(r1 * r2)
  }, error = function(e) { paste("Error with r1 * r2:", e$message) })
```

```{r}
Division_r1_r2 = tryCatch({
    show(r1 / r2)
  }, error = function(e) { paste("Error with r1 / r2:", e$message) })
```

```{r}
Addition_r1_r3 = tryCatch({
    show(r1 + r3)
  }, error = function(e) { paste("Error with r1 + r3:", e$message) })
```

```{r}
Multiplication_r1_r3 = tryCatch({
    show(r1 * r3)
  }, error = function(e) { paste("Error with r1 * r3:", e$message) })
```

```{r}
Division_r2_r3 = tryCatch({
    show(r2 / r3)
  }, error = function(e) { paste("Error with r2 / r3:", e$message) })
Division_r2_r3
```

```{r}
Quotient_r1 = tryCatch({
    quotient(r1)
  }, error = function(e) { paste("Error with quotient(r1):", e$message) })
```

```{r}
Quotient_r2 = tryCatch({
    quotient(r2)
  }, error = function(e) { paste("Error with quotient(r2):", e$message) })
```

```{r}
Quotient_r2_digits_3 = tryCatch({
    quotient(r2, digits = 3)
  }, error = function(e) { paste("Error with quotient(r2, digits = 3):", e$message) })
```

```{r}
Quotient_r2_digits_3.14 = tryCatch({
    quotient(r2, digits = 3.14)
  }, error = function(e) { paste("Error with quotient(r2, digits = 3.14):", e$message) })
```

```{r}
Quotient_r2_digits_avocado = tryCatch({
    quotient(r2, digits = "avocado")
  }, error = function(e) { paste("Error with quotient(r2, digits = 'avocado'):", e$message) })
```

```{r}
Quotient_assignment_q2 = tryCatch({
    q2 <- quotient(r2, digits = 3)
    q2
  }, error = function(e) { paste("Error with q2 assignment:", e$message) })
```

```{r}
Quotient_r3 = tryCatch({
    quotient(r3)
  }, error = function(e) { paste("Error with quotient(r3):", e$message) })
```

```{r}
Simplify_r1 = tryCatch({
    simplify(r1)
  }, error = function(e) { paste("Error with simplify(r1):", e$message) })
Simplify_r1
```

```{r}
Simplify_r2 = tryCatch({
    simplify(r2)
  }, error = function(e) { paste("Error with simplify(r2):", e$message) })
Simplify_r2
```


```{r}
Simplify_r3 = tryCatch({
    simplify(r3)
  }, error = function(e) { paste("Error with simplify(r3):", e$message) })
Simplify_r3
```

**C**
```{r}
# Attempt to create a Rational object with a zero denominator
tryCatch({
  invalid_rational <- rational(1, 0)
}, error = function(e) {
  cat("Error caught: ", e$message, "\n")
})

# Attempt to create a Rational object with non-numeric inputs
tryCatch({
  malformed_rational <- rational("a", "b")
}, error = function(e) {
  cat("Error caught: ", e$message, "\n")
})

# Attempt to create a Rational object without providing both numerator and denominator
tryCatch({
  missing_input_rational <- rational(1)
}, error = function(e) {
  cat("Error caught: ", e$message, "\n")
})
```



## Problem 2-plotly
```{r}
#import the csv
market <- read.csv("df_for_ml_improved_new_market.csv")
```

**a**Regenerate the plot. Based on the plot, we can observe that the sales proportion of the Painting category is declining. The Photograph category saw a significant increase in 1999, followed by a stable trend. The Print category started appearing in 2000, with a relatively small but steady share. The Sculpture category reached its lowest proportion in 1999, then increased and stabilized afterward.

```{r}
#get a new dataframe
#including years and the counts of items of different genres
genre_columns <- grep("Genre___", names(market), value = TRUE)
genre_data <- market %>%
  select(year, all_of(genre_columns)) %>%
  group_by(year) %>%
  summarise(across(everything(), sum))

#get the genre-percentage
genre_percentage_data <- genre_data %>%
  pivot_longer(cols = -year, names_to = "Genre", values_to = "Count") %>%
  group_by(year) %>%
  mutate(Percentage = Count / sum(Count) * 100) %>%
  mutate(Genre = gsub("Genre___", "", Genre))

# 使用 Plotly 创建堆叠条形图
fig <- plot_ly(
  genre_percentage_data,
  x = ~year,
  y = ~Percentage,
  color = ~Genre,
  type = "bar",
  text = ~paste0(round(Percentage, 1), "%"),
  textposition = "inside"
)

fig <- fig %>%
  layout(
    title = "Distribution of Genre of Art Sales Over Years",
    xaxis = list(title = "Year"),
    yaxis = list(title = "Percentage", ticksuffix = "%"),
    barmode = "stack",
    legend = list(title = list(text = "Genre"), orientation = "h")
  )

fig
```

**b**Observing the plot, we can see that the overall trend in art sales prices aligns closely with the trends in sales prices across different categories. The Photograph category is the primary influencing factor on the overall trend, while the price trend of Print artworks often contrasts with the overall price trend.

```{r}
#get a new dataframe
#including years and the counts of items of different genres
genre_columns <- grep("Genre___", names(market), value = TRUE)
sales_data <- market %>%
  select(year, price_usd, all_of(genre_columns)) %>%
  pivot_longer(cols = starts_with("Genre___"), names_to = "Genre", values_to = "Is_Genre") %>%
  filter(Is_Genre == 1) %>%
  mutate(Genre = gsub("Genre___", "", Genre)) %>%
  group_by(year, Genre) %>%
  #calculate average USD price based on genre
  summarise(Average_Price = mean(price_usd, na.rm = TRUE), .groups = "drop")

#create the figure
fig <- plot_ly(
  sales_data,
  x = ~year,
  y = ~Average_Price,
  color = ~Genre,
  type = "scatter",
  mode = "lines+markers",
  text = ~paste("Year:", year, "<br>Genre:", Genre, "<br>Average Price (USD):", round(Average_Price, 2))
)

fig <- fig %>%
  layout(
    title = "Changes in Sales Price Over Time by Genre",
    xaxis = list(title = "Year"),
    yaxis = list(title = "Average Price (USD)"),
    legend = list(title = list(text = "Genre"), orientation = "h")
  )

fig
```

## Problem 3-data.table

```{r}
#check the two dataframes
flights_dt <- as.data.table(flights)
airports_dt <- as.data.table(airports)
```

**a**Firstly, we get the table for departure-delay.
```{r}
# Calculate the mean and median departure delay for each airport.
dep_delay_table <- merge(
  flights_dt, 
  airports_dt, 
  by.x = "dest", 
  by.y = "faa", 
  all.x = FALSE, 
  all.y = FALSE
)[
  , .(
    mean_dep = mean(dep_delay, na.rm = TRUE),
    median_dep = median(dep_delay, na.rm = TRUE),
    flights = .N
  ),
  by = name
][flights >= 10  # Exclude airports with fewer than 10 flights.
][order(-mean_dep)  # Sort in descending order
][, .(name, mean_dep, median_dep)]

print(dep_delay_table)
```

Then, we get the table for arrival delay.
```{r}
# Calculate the mean and median arrival delay for each airport.
arr_delay_table <- merge(
  flights_dt, 
  airports_dt, 
  by.x = "dest", 
  by.y = "faa", 
  all.x = FALSE, 
  all.y = FALSE
)[
  , .(
    mean_arr = mean(arr_delay, na.rm = TRUE),
    median_arr = median(arr_delay, na.rm = TRUE),
    flights = .N
  ),
  by = name
][flights >= 10  # Exclude airports with fewer than 10 flights.
][order(-mean_arr)  # Sort in descending order
][, .(name, mean_arr, median_arr)]  

print(arr_delay_table)
```

**b**
```{r}
top_model <- flights_dt[planes_dt, # merge flights and planes on tailnum
                        on = .(tailnum)][, `:=`(
  time = air_time / 60, 
  # calculate the mph
  mph = distance / (air_time / 60)
)][, .(
  # calculate the avg_speed
  avg_speed = mean(mph, na.rm = TRUE),
  # the number of filghts
  numflights = .N
), 
by = model][order(-avg_speed)][1] #order table in descending average mph

print(top_model)
```
